---
title: "Social Network Analysis"
---

The following data is collected from an out of school time program for middle school girls (Digital Youth Divas) to engage in designing and making experiences in a f2f program. The program includes a story that is interwoven through the experiences. As part of the program, they spend time in a social learning network where they submit their work, look at other people's submitted work, develop profiles, get their work assessed, look at other people's profiles, comment, etc.

We will be using the "igraph" package to visualize different relationships. You can read more about igraph here:

http://igraph.org/r/

Upload the data files.

```{r}
NODE <- read.csv("dynDivas_Nodes.csv", sep = ",", header = TRUE)

EDGE <- read.csv("dynDYD_Edges.csv", sep = ",", header = TRUE)
EDGE <- EDGE[,c(4,5,1,2,3)]
```

iGraph requires data to be in a particular structure. There are several structures that it can use but we will be using a combination of an "edge list" and a "vertex list". As you might imagine the edge list contains a list of all the relationships between students and teachers and any characteristics of those edges that we might be interested in. There are two essential variables in the edge list a "from" variable and a "to" variable that descibe the relationships between vertices (a disciplinary action is given "from" a teacher "to" a student for example). 

The vertex list contains all the characteristics of those vertices.

Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#Load the igraph package

library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=NODE)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=NODE$role)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.size = 0.5, vertex.color=NODE$role, edge.color=EDGE$Action, edge.arrow.size=0.2, vertex.label=NA)

#Other attributes you may want to change are:

#Arrow size
edge.arrow.size=

#Remove the labels
vertex.label=NA

#Vertex size
vertex.size=
  
#And many others!
```

Play around with different graphs and then save your final version to the repo. Create a Zotero item for this R program and then commit, push and pull.
